---
title: "Fundamentals of R"
author: "R Trainer: Binod Jung Bogati (linkedin.com/in/bjungbogati)"
copyright: "© 2025. Unauthorized distribution or reuse prohibited."
format: html
editor: visual
---

## Running Your First Code

> R can be your calculator.

```{r}
1 + 5  

10 - 5  

10 * 5 

10 / 2 
```

## Your Turn 01

-   Add: 10 + 20
-   Substract: 10 - 20

```{r}



```

> R can do statistical calculation too.

```{r}

1:10

sum(1:1000) 

mean(1:10)

median(10:20)

```

## Your Turn 02

-   Sum: 10 to 20
-   Mean: 100 to 500

```{r}



```

## Running Your First Plot

```{r}
plot(cars)
```

## Ask for Help

```{r}
?cars
```

## Comments

-   Comments are helper text in R program
-   Written using `#` in the statement.
-   Makes our code understandable.

```{r}
1 + 5  # Add

10 - 5  # Subtract

10 * 5 # Multiply

10 / 2 # Divide
```

### Reserved words (Keywords)

Reserved words (or keywords) are a set of words that have special meaning.

-   It cannot be used as an identifier (variable name, function name etc.).
-   You can type ?reserved in R console to see list of reserved words in R.

```         
if, else, repeat, while, function, for, next, break,
TRUE, FALSE, NULL, Inf, NaN, NA, NA_integer_, NA_real_,
NA_complex_, NA_character_
```

```{r}
?reserved
```

### Essential Tips for Beginners

1.  *File Paths in R*

``` r
# ✅ Recommended style (forward slashes work everywhere)  - Windows, Mac or Linux
"D:/Clinical_Trial/Data/SDTM"

# ❌ Avoid (Windows backslashes need escaping)
"D:\Clinical_Trial\Data\SDTM"  # Error-prone
```

2.  Absolute Path vs Relative Path

``` r
"D:/Clinical_Trial/Data/SDTM/dm.xpt" # absolute / full path 

"./Data/SDTM/dm.xpt" # relative path to working directory
```

3.  Working Directory

```{r}
getwd() # current working directory

# setwd("D:/Clinical_Trial/Data/SDTM/") 
# set working directory is not used here since we already have active project
```

2.  **Keyboard Shortcuts**:

-   Run current line: `Ctrl+Enter` (Win) / `Cmd+Enter` (Mac)
-   Comment/uncomment: `Ctrl+Shift+C`
-   Auto-complete: `Tab`

3.  **Getting Help**:

```{r}
?mean     # Shortcut for help
```
